\section{Related Work}

Classical algorithmic stability studies how a learning rule changes when the
training sample is perturbed, primarily in order to control generalization under
an i.i.d.\ sampling model.  The canonical reference is Bousquet and Elisseeff,
who formalize several stability notions around replace-one perturbations and
fresh-test evaluation points \cite{bousquet2002stability}.  Kutin and Niyogi
expand this landscape with weaker and almost-everywhere variants
\cite{kutin2002almost}, while Feldman and Vondr\'ak and later Bousquet,
Klochkov, and Zhivotovskiy sharpen high-probability and rate-sensitive uniform
stability bounds \cite{feldman2019high,bousquet2020sharpening}.  Our paper does
not compete with this line on sharper generalization bounds.  Instead, we use
it as the baseline vocabulary against which a more microscopic perturbation
calculus can be compared.

\subsection{Algorithmic stability and generalization}

Most modern stability results are phrased around replacing one training example
and then evaluating expected or fresh-test loss
\cite{bousquet2002stability,kutin2002almost,mukherjee2006learning}.  This
perspective is powerful for generalization theory, but it can obscure a simpler
definition-level point relevant for local rules: ``change one sample point'' is
not a unique operation.  Delete-one, add-one, replace-one, and leave-one-out
share a family resemblance, yet they differ both in how the sample changes and
in where the learner is evaluated afterward.  Our contribution in this section
is therefore not a new stability hierarchy in the sense of
\cite{bousquet2002stability}, but a deterministic finite-sample reframing that
makes the perturbation operation and the evaluation point explicit.

\subsection{Nearest-neighbor classification and deleted estimates}

The nearest-neighbor line begins with Fix and Hodges and the classical risk
results of Cover and Hart, together with the consistency framework later
formalized by Stone \cite{fix1951discriminatory,cover1967nearest,cover1968estimation,stone1977consistent}.  Those works study statistical performance under
distributional sampling rather than worst-case single-perturbation effects.
Closer to our setting are deleted-estimate and local-rule analyses such as
Rogers and Wagner's finite-sample bound for local discrimination rules and the
distribution-free deleted-estimate inequalities of Devroye and Wagner
\cite{rogers1978finite,devroye1979distribution,devroye1979local}.  These are
important precedents because they show that leave-one-out style evaluation has a
substantial classical history around nearest-neighbor methods.  Even so, their
goal is not to separate LOO from adversarial replace-one at the level of fixed
finite metric witnesses with explicit deterministic tie-breaking.

\subsection{Leave-one-out, replace-one, and conformal stability certificates}

Kearns and Ron connect algorithmic stability with leave-one-out
cross-validation, already emphasizing that a CV-style perturbation has its own
logic \cite{kearns1999algorithmic}.  Recent conformal-prediction work makes
related distinctions operational again, often in service of coverage
certificates rather than 0--1 prediction loss.  Examples include transductive
stability analyses and training-conditional or stable conformal guarantees
\cite{ndiaye2022stability,liang2023learning,pournaderi2024training,lee2025framework}.  We cite this line not because our paper is a conformal-prediction
paper, but because it reinforces the message that leave-one-out style control
and replacement-based control can lead to different guarantees.  Our setting is
more concrete and more local: deterministic k-NN on finite metric spaces with
explicit top-$k$ neighbor ordering.

\subsection{Finite metric witnesses and local learning rules}

Metric-space analyses of k-NN typically appear in consistency or geometric
generality results rather than perturbation-separation examples
\cite{collins2020consistency}.  We did not identify, in the sources inspected
for this draft, a prior finite graph-metric example whose explicit purpose is to
show that fixed-sample LOO stability need not control fixed-sample adversarial
replace-one stability for deterministic 1-NN under the conventions used here.
This absence is not enough to justify a sweeping novelty claim.  It does,
however, support a careful contribution statement: the paper's distinctive value
lies in combining definition hygiene, explicit deterministic k-NN perturbation
analysis, and a reproducible finite witness/certificate pipeline.

\paragraph{What this paper does not claim.}
We do not claim a new generalization bound, a contradiction to classical
consistency results, or a theorem-level odd-$k$ lifting result from current
computational candidates alone.  The paper's main ambition is narrower and, we
think, cleaner: to make precise which one-sample perturbations are being
discussed, how they interact at the uniform level, and how sharply they can
separate at the pointwise finite-metric level for deterministic nearest
neighbors.
