\section{Uniform Perturbation Calculus}
\label{sec:uniform-calculus}

This section answers: what formal relationship holds between replace-one and
delete-one / insert-one at the uniform (worst-case) level?

\subsection{Replace-one as delete plus insert}

At the level of perturbation operations, replace-one is the composite move
obtained by deleting an occurrence and then inserting a new labeled point.
Concretely,
\[
S^{i \leftarrow z} = (S^{-i}) \oplus_i z.
\]
That observation is almost tautological at the sample level, but it yields the
correct direction of implication for worst-case stability control.

\begin{note}[Uniform insert-one stability]
When we speak of bounded insert-one stability $\beta_{\ins}(n)$ uniformly over
all insertion positions, we mean: for every ordered sample $T$ of size $m$,
every insertion position $j \in \{1,\ldots,m+1\}$, every labeled point $z$, and
every query-label pair $(x,y)$,
\[
\left|\ell(A(T),(x,y)) - \ell(A(T \oplus_j z),(x,y))\right| \le \beta_{\ins}(m).
\]
The bound may depend on the sample size $m$ before insertion.
\end{note}

\subsection{Uniform stability bound}

\begin{proposition}[Uniform replace-one via delete-and-insert]
\label{prop:4.1}
Fix a deterministic learning rule $A$ that maps ordered samples to predictors
and a bounded loss $\loss \in [0,1]$.  Suppose that:
\begin{enumerate}[leftmargin=*,label=(\arabic*)]
    \item Delete-one stability is bounded by $\beta_{\del}(n)$ uniformly over
        all ordered samples of size $n$, all indices, and all query-label
        pairs.
    \item Insert-one stability is bounded by $\beta_{\ins}(n)$ uniformly over
        all ordered samples of size $n$, all insertion positions $j$, all
        admissible labeled points $z$, and all query-label pairs.
\end{enumerate}
Then every replace-one perturbation satisfies
\[
\left|\loss(A(S),(x,y))-\loss(A(S^{i\leftarrow z}),(x,y))\right|
\le \beta_{\del}(n)+\beta_{\ins}(n-1).
\]
\end{proposition}

\begin{proof}
For fixed $S,i,z,(x,y)$, let $T=S^{-i}$.  By the insertion convention,
$S^{i\leftarrow z}=(S^{-i})\oplus_i z$.  Therefore
\begin{align*}
\left|\loss(A(S),(x,y))-\loss(A(S^{i\leftarrow z}),(x,y))\right|
&\le
\left|\loss(A(S),(x,y))-\loss(A(T),(x,y))\right| \\
&\quad +
\left|\loss(A(T),(x,y))-\loss(A(T\oplus_i z),(x,y))\right| \\
&\le \beta_{\del}(n) + \beta_{\ins}(n-1)
\end{align*}
by the triangle inequality and the uniform assumptions.  The last line uses
that $S^{-i}$ has size $n-1$, so the insert-one bound applies with that sample
size.
\end{proof}

\Cref{prop:4.1} is deliberately abstract.  It is not specific to $k$-NN, and it
does not by itself tell us whether the bound is informative for a particular
local learner.  Its role is to record the algebra of perturbation types before
we examine the combinatorics of deterministic nearest neighbors.

\subsection{Why LOO cannot substitute for delete-one control}

The proposition uses delete-one stability evaluated at the same arbitrary query
as the replacement perturbation.  LOO does not provide this, because its
evaluation point is tied to the deleted occurrence.  A pointwise LOO indicator
cannot simply be dropped into the role of delete-one control in
\Cref{prop:4.1}, because LOO changes the evaluation point along with the
sample.  Similarly, a fresh-test or expected-loss stability statement from the
classical literature does not automatically transfer to the worst-case
fixed-query indicators used in the present paper.  The whole point of the
calculus is to prevent these invalid transfers between notions.

The decomposition is sharp in a semantic sense: neither deletion-only evidence
nor LOO evidence can substitute for replacement control, as the separation
theorems of \Cref{sec:minimal-1nn,sec:odd-k-separation,sec:even-k-separation}
demonstrate.
