\section{Even-$k$ Separation and the Role of Tie-Breaking}
\label{sec:even-k-separation}

This section answers: does the separation extend to even $k$?  The answer is
yes, but with a crucial dependence on deterministic tie-breaking.

\subsection{Construction}

Let $k = 2m$ for an integer $m \ge 1$.  Use the same two-point metric space.
Define the ordered sample
\[
S_m^{\text{even}} \;=\; \big((a,0)_1,\ldots,(a,0)_m,\;
                            (a,1)_1,\ldots,(a,1)_m,\;
                            (b,0)_1,\ldots,(b,0)_m\big),
\]
of size $n = 3m$.  Fix $k = 2m$ and query-label pair $(a,0)$.

\subsection{Original prediction relies on tie-breaking}

At query $a$, the top-$k$ neighbors are $m$ copies of $(a,0)$ and $m$ copies of
$(a,1)$.  The signed margin is
\[
\M_k(S_m^{\text{even}}, a) = m(-1) + m(+1) = 0,
\]
and by the deterministic tie-breaking rule ($0 \prec 1$) the prediction is $0$,
so the loss at $(a,0)$ is $0$.  Unlike the odd-$k$ case, the prediction is not
determined by a strict majority; it is a tie that must be broken.

\subsection{Tie-dependent LOO stability}

\paragraph{LOO at an $(a,0)$ occurrence.}
After deletion, the top-$k'$ at $a$ consists of $(m-1)$ copies of $(a,0)$, $m$
copies of $(a,1)$, and the first copy of $(b,0)$.  The margin is
\[
\M_{k'} = (m-1)(-1) + m(+1) + 1(-1) = 0,
\]
the prediction remains $0$ by tie-breaking, and the loss at $(a,0)$ is
unchanged.  Hence $\Delta_{\mathrm{loo}} = 0$.

\paragraph{LOO at an $(a,1)$ occurrence.}
After deletion, the top-$k'$ at $a$ consists of $m$ copies of $(a,0)$,
$(m-1)$ copies of $(a,1)$, and the first copy of $(b,0)$.  The margin is
\[
\M_{k'} = m(-1) + (m-1)(+1) + 1(-1) = -2,
\]
the prediction is $0$.  The original loss at $(a,1)$ was $1$; after deletion
the loss remains $1$.  So $\Delta_{\mathrm{loo}} = 0$.

\paragraph{LOO at a $(b,0)$ occurrence.}
After deletion, the top-$k'$ at $b$ consists of $(m-1)$ copies of $(b,0)$, $m$
copies of $(a,0)$, and the first copy of $(a,1)$.  The margin is
\[
\M_{k'} = (m-1)(-1) + m(-1) + 1(+1) = -(2m-2),
\]
which is $\le 0$ for all $m \ge 1$.  The loss at $(b,0)$ is unchanged.
Therefore $\Delta_{\mathrm{loo}} = 0$.

\subsection{Replacement instability}

Replace the first occurrence $(a,0)_1$ with $(a,1)$.  The top-$k$ at $a$ now
contains one copy of $(a,1)$ at the top, $(m-1)$ copies of $(a,0)$, and $m$
copies of $(a,1)$.  The margin is
\[
\M_k = 1(+1) + (m-1)(-1) + m(+1) = 2 > 0,
\]
so the prediction flips to $1$.  The loss at $(a,0)$ goes from $0$ to $1$,
giving $\Delta_{\mathrm{rep}} = 1$.

\begin{theorem}[Even-$k$ LOO/replace-one separation]
\label{thm:even-k-separation}
For every even $k = 2m$ with $m \ge 1$, the sample $S_m^{\text{even}}$
satisfies $\Delta_{\mathrm{loo}}(S_m^{\text{even}}, i) = 0$ for every index
$i$, while there exists an index $j$ and a replacement point $z$ such that
$\Delta_{\mathrm{rep}}(S_m^{\text{even}}, j, z, a, 0) = 1$.
\end{theorem}

\subsection{Why this case is weaker than odd $k$}

The odd and even constructions share the same metric space and the same
replacement mechanism, but the margin analysis reveals a structural difference.

\begin{enumerate}[leftmargin=*,label=(\arabic*)]
\item \textbf{Margin at the prediction threshold.}  For odd $k$, the original
    margin is $-1$, strictly negative.  The prediction reflects a genuine (if
    minimal) majority.  For even $k$, the original margin is $0$, an exact tie;
    the prediction is determined solely by the tie-breaking convention.  In this
    sense, the even-$k$ separation is weaker: it depends on the deterministic
    tie-breaking rule to stabilize the original prediction against LOO
    perturbations.

\item \textbf{LOO stability margin.}  For odd $k$, after deleting an $(a,0)$
    occurrence, the margin remains $-1$ (non-zero, no tie).  For even $k$, the
    post-deletion margin becomes $0$ (a tie), and LOO stability relies on the
    same tie-breaking rule to keep the prediction at $0$.  If the tie-breaking
    rule were randomized, the even-$k$ prediction after deletion would be $0$ or
    $1$ with equal probability, potentially breaking the LOO stability guarantee.

\item \textbf{Replace-one effect.}  In both cases, the margin increases by
    exactly $2$ after the replacement (from $-1$ to $+1$ for odd $k$; from $0$
    to $+2$ for even $k$).  The increment of $2$ follows from flipping one
    label from $0$ to $1$ in the top-$k$: the contribution of that occurrence
    changes from $-1$ to $+1$.
\end{enumerate}

\subsection{Tie-free even-$k$ separation as an open problem}

Unlike the odd-$k$ construction, the even-$k$ separation depends on
deterministic tie-breaking.  Under randomized tie-breaking, the conclusion may
change.  Whether there exists, for any even $k \ge 2$, a finite metric sample
for deterministic $k$-NN in which the original prediction at the designated
query is determined by a strict majority (margin $\neq 0$) and the
LOO/replace-one separation holds remains an open problem.  The constructions in
this paper all use margin $0$ for even $k$, which ties the separation to
tie-breaking.
