\section{Stability Hierarchy}
\label{sec:stability-hierarchy}

This section collects what does and does not follow among the stability notions
studied in this paper.

\subsection{Established implications}

At the uniform (worst-case) level, \Cref{prop:4.1} establishes:
\[
\text{uniform delete-one + uniform insert-one}
\;\Rightarrow\;
\text{uniform replace-one}.
\]

At the pointwise level, the following trivial implication holds:
\[
\text{pointwise delete-one (all queries)}
\;\Rightarrow\;
\text{LOO}.
\]
This is immediate from definitions: LOO is delete-one evaluated at a specific
query (the deleted occurrence).  If delete-one is stable at all queries, it is
stable at that specific query.

\subsection{Established separations}

The central non-implication of this paper is:
\[
\text{pointwise LOO}
\;\not\Rightarrow\;
\text{pointwise replace-one}.
\]

This is demonstrated for every odd $k \ge 1$
(\Cref{thm:odd-k-separation} plus the $k=1$ witness of
\Cref{sec:minimal-1nn}) and for even $k$ under deterministic tie-breaking
(\Cref{thm:even-k-separation}).

Additional separations at the pointwise level:

\begin{itemize}
\item \textbf{Pointwise LOO $\not\Rightarrow$ pointwise delete-one (all
    queries).}  LOO stability at the deleted occurrence does not imply
    delete-one stability at a different query.  A counterexample is given in
    \Cref{app:hierarchy-counterexamples}.
\item \textbf{Pointwise replace-one $\not\Rightarrow$ pointwise LOO.}  A
    replace-one perturbation that changes the prediction at an arbitrary query
    does not guarantee the LOO indicator will be nonzero, due to the different
    evaluation quantification.
\end{itemize}

At the uniform level:
\begin{itemize}
\item \textbf{Expected replace-one $\not\Rightarrow$ worst-case replace-one.}
    Rare adversarial configurations can produce instability even when expected
    stability is high.
\item \textbf{Worst-case replace-one $\Rightarrow$ expected replace-one.}
    The supremum dominates the expectation.
\end{itemize}

\input{outputs/tables/stability_hierarchy.tex}

\subsection{Open relationships}

Several relationships remain open:
\begin{itemize}
\item The converse of the delete-plus-insert decomposition: does uniform
    replace-one imply uniform delete-one or uniform insert-one separately?
\item The comparability of uniform delete-one and uniform insert-one: is one
    always at least as large as the other?
\item The uniform LOO/uniform replace-one relationship: can a uniform version
    of the LOO bound be compared to replace-one without the evaluation-point
    mismatch?
\item Distributional analogues of the pointwise separations demonstrated here.
\end{itemize}

These open questions point to a richer stability taxonomy that goes beyond the
classical replace-one-centered framework.
