\section{Experimental Protocol}
\label{app:experimental-protocol}

This appendix specifies the full experimental configuration used in
\Cref{sec:experiments}.

\subsection{Software and reproducibility}

All experiments are implemented in Python.  Random seeds are recorded per trial.
The full source code is part of the supplementary material.

\subsection{Synthetic graph benchmark}

\begin{itemize}
\item \textbf{Graph generation.}  Erd\H{o}s--R\'enyi model $G(n,p)$ with
    $n \in \{2,3,4,5\}$ vertices and $p = 0.5$.  Graphs that are not connected
    are rejected.
\item \textbf{Metric.}  Shortest-path distance on the generated graph.
\item \textbf{Sample generation.}  For each graph, we draw a labeled sample of
    size $n \le 10$ with controlled:
    \begin{itemize}
        \item \emph{Duplicate ratio}: fraction of occurrences assigned to the
            same vertex as a previous occurrence.
        \item \emph{Conflict ratio}: fraction of vertices with occurrences of
            both labels.
        \item \emph{Label noise}: fraction of occurrences whose labels are
            randomly flipped after initial assignment.
    \end{itemize}
\item \textbf{Neighborhood sizes.}  $k \in \{1,3,5,7\}$.
\item \textbf{Trials.}  50 independent trials per configuration.
\item \textbf{Evaluation.}  For each trial, we compute the pointwise LOO and
    replace-one indicators for all indices, all admissible replacement points,
    and all query-label pairs in the metric space.
\end{itemize}

\subsection{UCI tabular benchmark}

\begin{itemize}
\item \textbf{Datasets.}  Iris (150 instances, 4 features), Wine (178
    instances, 13 features), Breast Cancer Wisconsin (569 instances, 30
    features), Digits (1797 instances, 64 features).
\item \textbf{Binary conversion.}  Class~0 vs.\ the rest.  For datasets with
    more than two classes, the first class (index~0) is treated as the positive
    class and all others are merged into class~1.
\item \textbf{Metric.}  Euclidean distance.
\item \textbf{Subsampling.}  For each dataset, 50 random subsamples of up to
    100 points are drawn without replacement.
\item \textbf{Neighborhood sizes.}  $k \in \{1,3,5,7\}$ (evaluated only when
    subsample size $\ge k$).
\item \textbf{Evaluation.}  Same as synthetic benchmark: compute LOO and
    replace-one indicators for all indices and queries.
\item \textbf{Confidence intervals.}  95\% Wilson score intervals reported
    where applicable.
\end{itemize}

\subsection{Diagnostic}

The margin-based diagnostic computes $|\M_k(S,x)|$ for each query $x$ in the
metric space.  A query is flagged as potentially vulnerable if
$|\M_k(S,x)| \le 2$.  For odd $k$, this condition is both necessary and
sufficient for the existence of a single replacement that flips the prediction
at $x$.
