\section{Complete Stability Hierarchy}
\label{sec:stability-hierarchy}

The stability notions introduced in \Cref{sec:definitions} and the separation
results developed above form part of a larger hierarchy of perturbation-stability
relationships. This section provides a systematic map, marking each
relationship as a theorem, a counterexample, or an open question.

\subsection{Hierarchy Table}

\input{outputs/tables/stability_hierarchy.tex}

\begin{proposition}[Pointwise delete-one and insert-one are incomparable]
\label{prop:del-not-impl-ins}
\label{prop:ins-not-impl-del}
At the pointwise level, delete-one stability does not imply insert-one
stability, and vice versa. Explicit counterexamples are given in the
appendix (\Cref{sec:proof-appendix}). Whether the same separation holds at
the uniform (worst-case) level remains an open problem.
\end{proposition}

\subsection{Key Relationships}

We highlight four essential relationships from the table.

\subsubsection{Uniform delete+insert $\Rightarrow$ uniform replace-one}

This is \Cref{prop:4.1}: the triangle inequality on losses gives
$\beta_{\mathrm{rep}}(n) \le \beta_{\mathrm{del}}(n) + \beta_{\mathrm{ins}}(n-1)$.
The converse does not follow from this bound and remains open at the uniform
level. At the pointwise level, the separation theorems below show that
replace-one stability can fail even when delete-one and insert-one are
individually stable, confirming that the decomposition is algebraically sharp.

\subsubsection{Pointwise LOO vs.\ pointwise replace-one}

This is the central separation of the paper. By \Cref{thm:odd-k-separation}
and \Cref{thm:even-k-separation}, pointwise LOO stability does not imply
pointwise replace-one stability, for both odd and even $k$. The converse
(pointwise replace-one $\Rightarrow$ pointwise LOO) is trivially false,
because replace-one stability is defined for a fixed query-label pair, while
LOO evaluates at the deleted occurrence itself.

A more abstract way to see the structural difference is the following.

\begin{proposition}[LOO only probes training-location-coupled perturbations]
\label{prop:loo-coupling}
For any deterministic learning rule, the LOO indicator
$\Delta_{\mathrm{loo}}(S,i)$ evaluates the loss at the deleted occurrence
$(x_i, y_i)$ itself. In contrast, the replace-one indicator
$\Delta_{\mathrm{rep}}(S,i,z,x,y)$ can evaluate any query-label pair
$(x,y)$, including those far from any training occurrence. Consequently, LOO
stability can be perfect ($\Delta_{\mathrm{loo}}(S,i)=0$ for all $i$) even
when there exist pairs $(x,y)$ at which a single replacement changes the
prediction, because LOO never tests those pairs.
\end{proposition}

This coupling between perturbation and evaluation point is not an incidental
feature of LOO; it is the defining property that distinguishes cross-validation
stability from adversarial replacement robustness.

\subsubsection{Pointwise delete-one (all queries) vs.\ pointwise LOO}

Pointwise delete-one is stronger than LOO: it requires stability at all
queries $(x,y)$, not just at the deleted occurrence. Therefore pointwise
delete-one stability implies LOO stability. The converse is false.

\begin{proposition}[LOO does not imply pointwise delete-one]
\label{prop:loo-not-impl-del}
There exists an ordered sample $S$ such that
$\Delta_{\mathrm{loo}}(S,i) = 0$ for all $i$ while
$\Delta_{\mathrm{del}}(S,i,x,y) = 1$ for some $i,x,y$ with
$(x,y) \neq (x_i,y_i)$.
\end{proposition}

\begin{proof}
Let the metric space be the three-point linear metric $\{0,1,2\}$ with
$d(0,1)=d(1,2)=1$ and $d(0,2)=2$. Fix $k=1$ and the sample
$S = ((0,0), (2,1), (0,0), (2,1))$. Every point appears with a same-label
duplicate, so for each index $i$, the $1$-NN prediction at the deleted
occurrence is unchanged after deletion: the duplicate ensures the prediction
remains the same. Hence $\Delta_{\mathrm{loo}}(S,i)=0$ for all $i$.

Now consider delete-one at index $0$ (the first $(0,0)$) with the query
$(1,0)$. The original $1$-NN at $1$ is a tie between $(0,0)$ (index $0$,
distance $1$) and $(2,1)$ (index $1$, distance $1$), resolved by sample-order
tie-breaking to $(0,0)$, predicting $0$; the loss at $(1,0)$ is $0$. After
deleting index $0$, the remaining sample is $((2,1), (0,0), (2,1))$, and the
$1$-NN at $1$ is again a tie between $(2,1)$ (index $0$, distance $1$) and
$(0,0)$ (index $1$, distance $1$), now resolved to $(2,1)$, predicting $1$.
Thus the loss at $(1,0)$ becomes $1$, giving
$\Delta_{\mathrm{del}}(S,0,1,0)=1$.
\end{proof}

\subsubsection{Deterministic tie-breaking vs.\ randomized tie-breaking}

Our tie-breaking convention ($0 \prec 1$) affects the stability landscape,
particularly for even $k$.

\begin{proposition}[Deterministic tie-breaking can mask instability]
\label{prop:tie-mask}
There exists a sample $S$ and even $k$ such that all LOO and delete-one
indicators vanish under deterministic tie-breaking ($0 \prec 1$), but a
randomized tie-breaking rule would reveal instability with positive
probability.
\end{proposition}

\begin{proof}
Take the even-$k$ sample from \Cref{thm:even-k-separation} with $k=2$:
$S = ((a,0), (a,1), (b,0))$. Under deterministic tie-breaking, the prediction
at $a$ is $0$ (tie goes to $0$) both before and after deleting $(a,0)$, so
the LOO indicator is $0$.

Under randomized tie-breaking, define the expected LOO indicator as
$\mathbb{E}[\Delta_{\mathrm{loo}}(S,i)]$ where the expectation is over
independent random tie-breaking coins. When the margin is $0$ (exact tie),
the prediction is $0$ or $1$ with equal probability $1/2$. The original
prediction at $a$ (margin $0$) is therefore $0$ or $1$ with probability
$1/2$ each. After deleting $(a,0)$, the margin is again $0$, and the
post-deletion prediction is independent $0$ or $1$ with probability $1/2$
each. The expected loss difference is
$\mathbb{E}[|\ell_{\text{orig}} - \ell_{\text{post}}|] = 1/2 \neq 0$, so the
expected LOO indicator is non-zero under independent randomized tie-breaking.
\end{proof}

This shows that deterministic tie-breaking can hide instabilities that would
be visible under randomization. The odd-$k$ separation
(\Cref{thm:odd-k-separation}) does {\em not} rely on this masking effect,
since the margins are strictly non-zero throughout.

\subsection{Summary Diagram}

The hierarchy can be summarized by the following diagram (solid arrows indicate
established implications; dashed arrows indicate open relationships):

\[
\begin{array}{c}
\text{Uniform delete+insert} \\
\downarrow \\
\text{Uniform replace-one} \\
\downarrow \\
\text{Pointwise delete-one (all queries)} \\
\downarrow \\
\text{Pointwise LOO} \qquad \text{Pointwise replace-one (fixed query)} \\
\searrow \qquad \swarrow \\
\text{[Separated: \Cref{thm:odd-k-separation}, \Cref{thm:even-k-separation}]}
\end{array}
\]

The key negative result -- the separation between LOO and replace-one -- is
the main message of this paper. Further relationships (uniform delete-one vs.\
insert-one, expected vs.\ worst-case, fresh-test vs.\ fixed-query) are
recorded in the table above; some remain open at the uniform level.

\input{outputs/tables/hierarchy_summary.tex}
